Light inextensible strings under tension in the Schwarzschild geometry

### 2018-10-01 string_microarticle.tex
tidyup and terseification

%% main text
\section{Introduction}
\label{intro}
In this short paper I consider the behaviour of a light inextensible
string in the exterior Schwarzschild metric.  A light string is not
subject to gravity, and being under tension means that its
configuration is an extremum of proper length.  Of course, no real
string is perfectly weightless or inextensible.  With tension~$T$,
gravitational acceleration $g$ and (proper) mass per (proper) unit
length~$\mu$, then ideal behaviour will be appropriate for
lengthscales $L\lesssim\frac{T}{\mu g}$.  Using Newtonian expressions
for the strength of gravity at the event horizon gives
$L\ll\frac{TGM}{\mu c^2}$.  For the string to behave itself near the
event horizon on lengthscales comparable with the Schwarzschild radius
would require $T\gtrsim\mu c^2$.  Exceeding this criterion would
violate the dominant energy condition (and in this context we note
that transverse waves would propagate at speed~$\sqrt{T/\mu}\sim c$).

Nevertheless, the idealization of a light inextensible string is an
interesting and informative limit: its analysis interprets the concept
``straight line'' and suggests a possible way of manipulating a black
hole.  The Schwarzschild metric

@@

where units in which~$2GM=c=1$ are used, has been known for over a
century and makes a number of counter-intuitive predictions.  Here I
consider the spatial component~$ds^2= \frac{dr^2}{1-1/r} +
r^2d\phi^2$, confined to equatorial points~($\phi=\pi/2$); we
understand~$1<r<\infty$ throughout.  A string of constant~$r$ will,
trivially, have proper length~$2\pi r$.

@@

\begin{equation}\label{radial_string_length}
  \int_{r=r_1}^{r_2}\frac{dr}{\sqrt{1-2M/r}}=
  \left.
  \sqrt{r(r-1)} +\frac{1}{2}\log\left(
  \frac{r+\sqrt{r-1}}{r-\sqrt{r-1}}\right)
  \right|_{r_1}^{r_2}
  \end{equation}

@@

\subsection{Radial and angular components}

@@

\begin{equation}
%  \int_{p_1}^{p_2}\sqrt{\frac{\left(r'\right)^2}{1-2M/r} + r^2}d\phi=
  \int_{p_1}^{p_2}\left(\frac{\left(r'\right)^2}{1-2M/r} + r^2\right)^\frac{1}{2}d\phi=
  \int_{p_1}^{p_2}F\left(r,r'\right)d\phi.
\end{equation}

@@

Figure~\ref{closest_approach_non_self_intersecting} shows
nonselfintersecting strings indexed by distance of approach to the
singularity, which occurs at~$\phi=0$.  Strings with a sufficiently
close passage to the event horizon self-intersect; numerical methods
show that the critical value is~$\simeq 1.076109317$, with an
uncertainty of 1 in the last digit: if a string with two free ends
passes closer than this, it forms a closed loop.
Figure~\ref{closest_approach_non_self_intersecting} shows some
non-self-intersecting strings and
figure~\ref{closest_approach_self_intersecting} shows some strings
that self-intersect.

@@

Figure~\ref{one_free_end_r_equals_2} shows some strings that pass
through~$\left(2,0\right)$ and tangent to the event horizon.

\subsection{Strings with one free end}

@@

This work considered only the Schwarzschild metric and generalizing it
to the Kerr metric~\cite{asdfsadfasfd} might be interesting.

>>>>

%% main text
\section{Introduction}
\label{intro}
In this short paper I consider the behaviour of a light inextensible
string in the exterior Schwarzschild metric.  A light string is not
subject to gravity, and being under tension means that its
configuration is an extremum of proper length.  Of course, no real
string is perfectly weightless or inextensible.  With tension~$T$,
gravitational acceleration $g$ and (proper) mass per (proper) unit
length~$\mu$, then ideal behaviour will be appropriate for
lengthscales $L\lesssim\frac{T}{\mu g}$.  Using Newtonian expressions
for the strength of gravity at the event horizon gives
$L\ll\frac{TGM}{\mu c^2}$.  For the string to behave as though light
on lengthscales comparable with the Schwarzschild radius would require
$T\gtrsim\mu c^2$.  Exceeding this criterion would violate the
dominant energy condition (and in this context we note that transverse
waves would propagate at speed~$\sqrt{T/\mu}\sim c$).  Nevertheless,
the idealization of a light inextensible string is an interesting and
informative limit: its analysis interprets the concept ``straight
line'' and suggests a possible way of manipulating a black hole.  The
Schwarzschild metric

@@

where units in which~$2GM=c=1$ are used, has been known for over a
century~\cite{schwarzschild1916} and makes a number of
counter-intuitive predictions.  Here I consider the spatial
component~$ds^2= \frac{dr^2}{1-1/r} + r^2d\phi^2$, confined to
equatorial points~($\phi=\pi/2$); we understand~$1<r<\infty$
throughout.  A string of constant~$r$ will, trivially, have proper
length~$2\pi r$.

@@

\begin{equation}\label{radial_string_length}
  \int_{r=r_1}^{r_2}\frac{dr}{\sqrt{1-1/r}}=
  \left.
  \sqrt{r(r-1)} +\frac{1}{2}\log\left(
  \frac{r+\sqrt{r-1}}{r-\sqrt{r-1}}\right)
  \right|_{r_1}^{r_2}
  \end{equation}

@@

@@

\begin{equation}
%  \int_{p_1}^{p_2}\sqrt{\frac{\left(r'\right)^2}{1-2M/r} + r^2}d\phi=
  \int_{p_1}^{p_2}\left(\frac{\left(r'\right)^2}{1-1/r} + r^2\right)^\frac{1}{2}d\phi=
  \int_{p_1}^{p_2}F\left(r,r'\right)d\phi.
\end{equation}

@@

Figure~\ref{closest_approach_non_self_intersecting} shows
nonselfintersecting strings indexed by distance of approach to the
singularity, which occurs at~$\phi=0$.  Strings with a sufficiently
close passage to the event horizon self-intersect; numerical methods
show that the critical value is~$\simeq 1.076109317$, with an
uncertainty of 1 in the last digit: if a string with two free ends
passes closer than this, it forms a closed loop.
Figure~\ref{closest_approach_non_self_intersecting} shows some
non-self-intersecting strings and
figure~\ref{closest_approach_self_intersecting} shows some strings
that self-intersect.  Figure~\ref{one_free_end} shows some strings
that pass through~$\left(2,0\right)$ and tangent to the event horizon.

@@

@@

This work considered only the Schwarzschild metric and generalizing it
to the Kerr metric~\cite{kerr1963} might be interesting.
